\textls[-0]{\thetitle}

### 2023-03-05 canonicitymodel.tex
Working on appendices.

We follow the reducibility argument from~\citet{coquand2018canonicity,sterling2019algebraic, kaposi2019gluing} to construct the canonicity model. We will first specify the formulation of the most basic canonicity theorem.

@@

Now we sketch out the big picture of the proof. The canonicity model $"Tm"^C~\Gamma^C~"T"^C$ will map each syntax piece $("t : Tm Γ T") \mapsto "(t, tₚ) :" "Tm"^C~\Gamma^C~"T"^C = \sum~("t : Tm Γ T")."Tm"_p~\Gamma^C~"T"^C~"t"$ to a dependent pair\footnote{This $"Tm"_p$ is usually called dependent model.}, of which the first part will be the same as the input syntax. Using this mapping on closed boolean term $"t : Tm ⋅ "\mathbb{B}$ will give us a proof of $"Tm"_p~\cdot^C~\mathbb{B}^C~"t"$, which will unfolds to our final goal $"t" \equiv "tt" + "t" \equiv "ff"$. Roughly, this can read as ``every closed boolean term is reducible''.  Thus once we can ``propagate'' $(\cdot)^C$ into every syntactic piece, and construct this model, we are done. In our (pseudo) Agda formulation, we use $(\cdot)_2$ to replace the notation of $(\cdot)^C$ because we consider most parts of this model are actually a dependent pair.

Another helpful analogy of the canonicity model is that $"Tm"_p~\Gamma^C~"T"^C~"t"$ can be understood as ``a reducible set'': $"t" \in \{$ Reducible Terms of type $T$ (and $T_p$) in context $\Gamma$ (and $\Gamma_p$) $\}$. The reformulation can be understood as an unfolding of the \textit{type-theoretic logical relation proof}.\footnote{We usually use predicate to encode ``belong to''/``is element of'' relation in type theory} If we use the terminology of the conventional formulation, this reducible set means if we apply its element $t$ with a substitution $\gamma$ from the set $\Gamma$ where we have $\Gamma_p$ witnesses the ``reducibility'' of $\Gamma$, then we will have $t[\gamma] \in T_p$ as a closed and \textbf{reducible} term. Now if $\Gamma$ is actually empty $\cdot$ and $T$ is $\cB$, we are done.

% $\Gamma_p$ witnesses the set of substitution $ \gamma \in \Gamma$ makes applied term $t[\gamma]$ closed and \textbf{reducible} of types $T_p$. Thus, we proved $t[\gamma] \in T_p$---the set of closed reducible term of type $T$---when $\gamma \in \Gamma_p$.

>>>>

We follow the reducibility argument from~\cite{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} to construct the canonicity model.
Each of their proof is based on QIIT, Category with Family\cite{dybjer1995internal} and Generalized Algebraic Theory \cite{cartmell1986generalised} as the syntax. However, our proof will still work directly on the earlier introdcued syntax for readability.
We will first specify the formulation of the most basic canonicity theorem.

@@

Like most logical relation proof,  we interpret each judgement and prove compatibility lemma.

### 2023-03-05 metatheory2.tex
Working on appendices.

In this section, we start by introducing the complete formulation of the syntax of \TT, 
and then we detail the proofs for syntactic translation (that translate away linkages), 
consistency and canonicity.

\subsection{Warm Up: Review Declarative MLTT}
We begin with a review of MLTT as \TT can be roughly considered as MLTT with linkages.
We work in an intrinsic-typed setting, so there is no un-typed preterms.

>>>>

In this section, we start by introducing the complete formulation of the syntax of \TT, 
and then we detail the models and proofs for syntactic translation (that translate away linkages), 
consistency and canonicity.

\subsection{Warm Up: Review Declarative MLTT}
We begin with a review of MLTT as \TT can be roughly considered as MLTT with linkages.
For our syntax, we work in an intrinsic-typed setting, so there is no un-typed preterms 
(i.e. every terms are well-typed by construction). We introduce judgemental equality $\equiv$
for derivations (of each judgements). In other words, we will have judgemental equality between terms, 
contexts and explicits substitutions. 
We will make sure all the introduced models and proofs respecting judgemental equality.

### 2023-03-05 standardmodel.tex
Working on appendices.

The big picture of the proof is that: (1) we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; (2) we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma. The main goal is to interpret $\denotesS{\bot} = \emptyset$ which implies it is impossible $\goodTermS{\cdot}{t}{\bot}$ and thus $\bot$ is syntactically non-derivable. We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$ and that $\denotesS{}$ respecting judgemental equality.\footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo) Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}

>>>>

The big picture of the proof is that: (1) we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; (2) we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma. The main goal is to interpret $\denotesS{\bot} = \emptyset$ which implies it is impossible $\goodTermS{\cdot}{t}{\bot}$ and thus $\bot$ is syntactically non-derivable. We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$ and that $\denotesS{}$ respecting judgemental equality.
% \footnote{Actually our 1 is acting exactly as an interpretation function, and in our (pseudo) Agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}
